---
title: "ANOVA By Hand"
author: "Jahiya Clark"
date: "11/28/2021"
output: html_document
---

# Question 1 (anova by hand)
```{r}

require(here)
rope <- read.csv(here("data", "rope.csv"))

rope$rope.type = factor(rope$rope.type)
levels(rope$rope.type)
                        
n_obs = nrow(rope)
n_groups = length(levels(rope$rope.type))             

ss_tot = sum((rope$p.cut - (mean(rope$p.cut)))^2)
df_tot = n_obs - 1

aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type), 
  FUN = mean)

aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) mean(x))

agg_resids = 
  aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) mean(x))

str(agg_resids)

agg_sq_resids = 
  aggregate(
    x = rope$p.cut,
    by = list(rope$rope.type),
    FUN = function(x) sum((x - mean(x))^2) )

str(agg_sq_resids)

ss_within = sum(agg_sq_resids$x)
df_within <- n_obs - n_groups

ss_among = ss_tot - ss_within
df_among <- n_groups -1

ms_among  =  ss_among / (n_groups - 1)
ms_within = ss_within / (n_obs - n_groups)
                  
f_ratio = ms_among / ms_within
f_ratio
f_pval = 1 - pf(f_ratio, df_among, df_within)
f_pval

```

# Question 1 (self-check)

```{r}

digits_check = 5

fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

anova_fit_1$Df == c(df_among, df_within)

round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)

```


# Question 2

Yes, I beleive the variances from the partioning variance: within-group are about equal, the values seem to be around 0.2, just looking like at it. 

# Question 3

The p-value of the Bartlett test is 0.00143, meaning we can reject the null hypothesis that variances are the same. 

```{r}
bartlett.test(p.cut ~ rope.type, data = rope)
```

# Question 4

Yes, I would say an ANOVA analysis is needed because just based on the graphical depiction of the raw data the variances seem equal but the Bartlett test had such a small p-value that we can argue the variances are quite different. There might be more difference that a statistical analysis can give us more information.

# Question 5

Blaze rope type is the base case. 

# Question 6

The mean percent cut of the base case rope is 0.36714

```{r}
fit_rope_1 = lm(p.cut ~ rope.type, data = rope)
summary(fit_rope_1)

fit_rope_1$coefficient[[1]]


```

# Question 7

The mean percent cut of the XTC rope type is 0.2655. 

```{r}

fit_rope_1$coefficient[[1]] + fit_rope_1$coefficient[[6]]
```



