---
title: "Lab 9"
author: "Jahiya Clark"
date: "11/13/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  code_folding: show
---
# Brown Creeper Chi- Square Test

```{r}
require(here)
birds   = read.csv(here("data", "bird.sta.csv"))
hab     = read.csv(here("data", "hab.sta.csv"))
birdhab = merge(birds, hab, by=c("basin", "sub", "sta"))
br_creeper_table = table(birdhab$s.edge, birdhab$BRCR > 0)[, 2:1]

chisq.test(br_creeper_table)
```

# Question 1

There is no significant relationship between brown creeper presence or absence and forest interior or forest edge. 

# Question 2 

My chi square value is 23.3, which is high, and my p value is 1.386e^-6, which is less than significance value of 0.05. Thus I can reject the null hypothesis. The two variables are dependent and have a relationship. I do believe brown creepers show a habitat preference for the interior of forest after also doing prop.test.

```{r}
prop.test(c(29, 314), c(173, 873), 
          alternative = c("less"))
```
# Building Models for ANOVA

```{r}
require(palmerpenguins)
fit_fl_sp = 
  lm(
    formula = flipper_length_mm ~ species,
    data = penguins)
```

# Question 3


```{r}

fit_species <- 
  lm(
    formula = penguins$body_mass_g ~ penguins$species
  )
```

# Question 4
```{r}
fit_sex <-  
  lm(
    formula = penguins$body_mass_g ~ penguins$sex)

```

# Question 5

```{r}
fit_both <- 
  lm(
    formula = penguins$body_mass_g ~ penguins$species * penguins$sex
  )
```


# Question 6

```{r}
boxplot(
  formula(fit_species),
  main = "Penguins by Body Mass",
  ylab = "Body Mass (g)",
  xlab = "",
  col = "pink"
)
```

# Question 7

```{r}
boxplot(
  formula(fit_sex),
  main = "Penguin's Sex by their Body Mass",
  ylab = "Body Mass (g)",
  xlab = "",
  col = "#C39BD3"
)
```

# Question 8

```{r}
boxplot(
  formula(fit_both),
  ylab = "Body Mass (g)",
  xlab = "",
  names = c("Female \nAdelie", "Female \nChinstrap", 
            "Female \nGentoo", "Male \nAdelie", "Male \nChinstrap",
            "Male \nGentoo"),las= 2,
  col = rainbow(6)
)
```

# Question 9

I think the fit_species model will have a problem with the homogeneity assumption because the Adelie and Chinstrap boxes are very similar. Additionally, because of that my model fit_both might have a problem as well. 

# Question 10

The null hypothesis of a Bartlett test is that the sample variances are the same. 

# Question 11

The p-value is 0.0500
```{r}
bartlett.test(penguins$body_mass_g ~ penguins$species)

```
# Question 12

The p-value is 0.0319

```{r}
bartlett.test(penguins$body_mass_g ~ penguins$sex)

```

# Question 13

The p-value of from the Bartlett test of homogeneity for observations of body mass (g) grouped by sex was 0.03194 and  species was 0.05005. 

```{r}
sex_group = aggregate(
  body_mass_g ~ sex,
  data = penguins,
  FUN = c)

sp_group = aggregate(
  body_mass_g ~ species,
  data = penguins,
  FUN = c)

bartlett.test(sex_group$body_mass_g)

bartlett.test(sp_group$body_mass_g)

```

# Question 14

After conducting the Bartlett test for homogeneity, I found some issue with the test comparing body mass by species. Because the p-value is 0.05005 I will not reject the null hypothesis. The cutoff of 0.05 is to me not equal to or less than but just less than 0.05 to be significant. 




