---
title: 'Final Project: R Reference'
author: "Jahiya Clark"
date: "12/2/2021"
output: html_document
---

# Part 1: R Reference Guide {.tabset .tabset-pills}


## Loading Data Packages 

```{r}
require(here)
library(palmerpenguins)
ginkgo <- data.frame(read.csv(here("data", "ginkgo_data_2021.csv")))
head(ginkgo)
```


## Data Structures

### c()

The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
  + I can’t combine character and numeric types in the same call to c()
Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)
```

### length()

The function length() can be used to get the length of vectors (including lists) and factors.
```{r}
length(ginkgo$ï..site_id)
```
### Matrix()

This function creates matrix from given data values. 
```{r}
matrix(data = 0:50, nrow = 5, ncol = 5, byrow = FALSE)
```
### Data.Frame()

The function data frame() creates a collection of variables related to each other in matrices and lists.

```{r}
head(data.frame(ginkgo))

```
```{r}
nrow(ginkgo)
ncol(ginkgo)
dim(ginkgo)
```


## Subsetting

### $
```{r}
head(ginkgo$seeds_present)
```

### []

```{r}
ginkgo[1,]

ginkgo[2,3]

head(ginkgo[,3])
```
### Subset()

```{r}
head(subset(penguins, species == "Adelie"))
```


## Numerical Data Exploration

### Summary()

The function summary() is  used to produce result summaries of the results of various model fitting functions. 
```{r}
summary(ginkgo)
```
### Mean()

The function mean() is used to get the result of the data's arithmetic mean. 
```{r}
mean(ginkgo$max_width)
```
### SD()

The function sd() is used to get the standard deviation of the interested data.

```{r}
sd(ginkgo$max_width)
```



## Graphical Data Exploration

### Scatterplot
```{r}
plot(max_width ~ max_depth, data = ginkgo, 
     xlab = "Max Leaf Depth", ylab = "Max Lead Width",
     main = "Scatterplot of Ginkgo Leaf Max Depth by Max Width",
     col = ginkgo$ï..site_id, xlim = c(5, 150), ylim = c(5, 150), pch= 5, cex = 1.5)
```


### Histogram

```{r}
require(palmerpenguins)
penguins <- data.frame(penguins)

hist(penguins$flipper_length_mm, breaks = 12, main = "Histogram of Penguin Flipper Lengths", xlab = "Flipper Length (mm)", col = cm.colors(12))
```

### Boxplot

```{r}
boxplot(ginkgo$petiole_length, xlab = "Petiole Length", main = "Boxplot of Petiole Length", col = "aliceblue")
```
```{r}
boxplot(ginkgo$seeds_present, ginkgo$max_depth, names = c("No Seeds", "Seed Present"), main = "Boxplot of Seed Presence 
        by the Max Leaf Depth", col = topo.colors(2))
```

### Par() and Mrow of Histograms

```{r}
par(mfrow = c(2, 2))

hist(ginkgo$max_width, main = "Histogram of Ginkgo Max Leaf Width", xlab = "Max Leaf Width", col = heat.colors(10))
hist(penguins$bill_length_mm, main = "Histogram of Penguin Bill Lengths", xlab = "Bill lengths (mm)", col= terrain.colors(13))

seat <- data.frame(Seatbelts)
hist(seat$DriversKilled, main = "Histogram of Drivers Killed from 
     Road Casualties in Great Britain
              1969-84", xlab = "Monthly number of Drivers Killed", 
    col =hsv(0.5, seq(0,.50,length.out = 7), .85) )
hist(iris$Petal.Length, main = "Histogram of Iris Petal Length", xlab = "Petal Length", col=c(2,4,6,8))
```


## Distribution Functions

### dnorm()

dnorm() calculates the probability density of normal distribution.

```{r}
dnorm(x = -8:8, mean = 1, sd = 3) 
```

### pnorm()

pnorm outputs the cumulative distribution of the normal distribution. 

```{r}
pnorm(0, mean = 0, sd = 1)
```

### qnorm()

qnorm outputs the quantile Q for any probability. 

```{r}
qnorm(0.17, mean = 0, sd = 1)
```

### dbinom()

dbinom() calculates the probability density of binomial distribution

```{r}
dbinom(x = 1:10, size = 10, prob = 0.4)
```

### pbinom()

pbinom  will give you the binomial distribution of an event.

```{r}
pbinom(69, size = 10, prob = 0.3)
```

### qbinom()

qbinom will output the binomial quantile function. 

```{r}
qbinom(p = 0.67, size = 5, prob = 0.47)
```


